---
title: "DSC1107 Formative Assessment 1"
author: "DESQUITADO, Yuichi P."
date: "2024-02-16"
output: html_document
---

Load the needed packages.

```{r}
library(tidyverse)
library(DescTools)
diamonds
```

# **Lecture 2**

Discrete Variables - We usually use bar charts to visualize discrete variables.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut),
           fill = "lightgreen",
           alpha = 0.5)
```

INTERPRETATION: This plot implies that 20000 diamonds have ideal cut, which is the greatest count compared to others, such as Premium, Very good, Good, and Fair.


Continuous Variables - We usually use histograms for continuous variable visualization.

```{r}
ggplot(data = diamonds) +
  geom_histogram(
    mapping = aes(x = carat),
    binwidth = 0.01,
    fill = "purple") +
  coord_cartesian(xlim = c(0, 3))

Mode(diamonds$carat)
```

INTERPRETATION: This plot shows that a value between 0.25 and 0.50 has the highest count in carat. If we check the mode of the dataset diamonds$carat, it is "0.3" with 2604 entries.

Discrete vs. Discrete - We use count plots to visualize covariation between two discrete variables.

```{r}
ggplot(data = diamonds) + 
  geom_count(mapping = aes(x = cut,
                           y = clarity),
             color = "blue")
```

INTERPRETATION: According to the plot, the most abundant diamonds with specific cut and clarity are ideal-VS2, which has around 5000 count.

Continuous vs. Continuous (Many to One) - Scatter plot is used for comparing covariations of continuous variables (many to one).

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = x,
                           y = y),
             color = "gold") +
  geom_abline(slope = 1,
              intercept = 0, 
              color = "red") + 
  coord_cartesian(xlim = c(3.5, 10), 
                  ylim = c(3.5, 10))
```

INTERPRETATION: Data being huddled around the centerline means that the data follows normal distribution.

Continuous vs. Continuous (One to One) - Line plot is used for comparing covariations of continuous variables (one to one).

```{r}
avgPriceByCarat <- diamonds %>%
  mutate(carat = round(carat, 1)) %>%
  group_by(carat) %>%
  summarise(avg_price = mean(price))
avgPriceByCarat

ggplot(data = avgPriceByCarat, aes(x = carat,
                                   y = avg_price)) + 
  geom_line(mapping = aes(x = carat,
                          y = avg_price),
            color = "magenta",
            linewidth = 1) +
  geom_area(fill = "pink",
            alpha = 0.5)
```

INTERPRETATION: This plot shows that the higher the carat, the higher the average price is, but there is a significant number of data that shows in between 3 and 4 carat, prices drop down almost the same to how 1-2 carats were priced.

Continuous vs. Discrete - We use boxplot to visualize the covariation between continuous and discrete variables.

```{r}
ggplot(data = diamonds,
       aes(x = cut,
           y = price,
           fill = cut,
           color = cut,
           alpha = 0.75)) +
  geom_boxplot(mapping = aes(x = cut,
                             y = price),
               outlier.color = "black")
```

INTERPRETATION: There is a inverse correlation between price and cut, wherein the more that the cut is getting higher in quality, the more its median price becoming lower and cheaper.

---

# **Lecture 3**

Exercise: Filter diamonds to those with ideal cut and at least 3 carats. How many such diamonds are there?

```{r}
filter(diamonds, cut %in% c("Ideal") & carat >= 3)
```

ANSWER: There are four diamonds that are in Ideal cut and has a carat greater than or equal to 3.

Exercise: Select all columns except x, y, z.

```{r}
select(diamonds, -x, -y, -z)
```

Exercise: Arrange diamonds in decreasing order of their length. How long is the longest diamond?

```{r}
arrange(diamonds, desc(depth))
```

ANSWER: Longest is 79.0.

Exercise: Add a variable called good_color that is TRUE if the color is D, E, F, G and FALSE otherwise.

```{r}
mutate(diamonds, good_color =
         if_else(
           condition = color %in% c("D", "E", "F", "G"),
           true = TRUE,
           false = FALSE))
```

Exercise: Use summarise to determine if there are any diamonds of at least one carat that cost less that $1000.

```{r}
summarise(diamonds, cheapDiamonds = any(price < 1000 & carat >= 1))
```

ANSWER: False (None)

Exercise: Compute the mean price for diamonds of volume at least one carat

```{r}
diamonds %>%
  filter(carat >= 1) %>%
  mutate(meanPrice = mean(price)) #%>%
  #ggplot() + 
  #geom_histogram(aes(x = meanPrice))
```

Exercise: Reproduce the output of count(diamonds, cut) via group_by() and summarise()

```{r}
count(diamonds, cut)
```
ANSWER:

```{r}
diamonds %>%
  group_by(cut) %>%
  summarise(n = n())
```

------

# **Lecture 4**

heights2 = read_csv(file = "heights2.csv") 

```{r}
table1
table2
table3
```
ANSWER: table1 utilized columns to express variables/observations per country and year, table2 utilized rows in separating the observed variable, which is count on the last column, while table 3 used ration of case/population to simplify the presentation to three columns, yet still separated by year.

Exercise: Use pivot_longer() to tidy table4b in a similar fashion. What is the difference between the code
used to tidy table4a and table4b?

```{r}
table4b
```

ANSWER: The table4a tackles cases, while table4b tackles population.

```{r}
table4b %>%
  pivot_longer(cols = c(`1999`, `2000`), names_to = "year", values_to = "population")
```

Exercises:

1. Why does this code fail?

ANSWER: Because the ``cols = c(1999, 2000)`` doesn't have backticks.

2. 

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes", NA, 10,
  "no", 20, 12
)
preg
```

```{r}
preg %>%
  pivot_longer(cols = c(male, female), names_to = "gender", values_to = "Answer")
```

